
# Integration

## Time series plot of satellite & sentiment metrics

Let's now integrate data from satellite and sentiment into the same monthly time series plot, first rescaling all metrics -1 to 1 for comparison.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

```{r}
library(here)
library(glue)
library(readr)
library(dplyr)
library(lubridate)
library(scales)
library(ggplot2)
library(plotly)

# paths
loc_name       <- "delhi"
satellite_csv  <- here(glue("data/no2_{loc_name}_ee-chart.csv"))
sentiments_csv <- here(glue("data/sentiments_{loc_name}.csv"))
metrics_csv    <- here(glue("data/metrics_{loc_name}.csv"))

if (!file.exists(metrics_csv)){
  satellite  <- read_csv(satellite_csv) %>% 
    transmute(
      date    = mdy(`system:time_start`),
      date_ym = floor_date(date, unit = "month"),
      no2     = NO2_column_number_density) %>% 
    filter(!is.na(no2)) %>% 
    group_by(date_ym) %>% 
    summarize(
      no2 = mean(no2, na.rm=T)) %>% 
    mutate(
      metric = "NO2",
      value  = rescale(no2, to = c(-1, 1))) %>% 
    select(metric, date_ym, value)
  
  sentiments <- read_csv(sentiments_csv) %>% 
    rename(date_ym = ym_date) %>% 
    mutate(
      metric = glue("{query_name}"),
      value  = rescale(tf_score, to = c(-1, 1))) %>% 
    select(metric, date_ym, value)
  
  bind_rows(
    satellite,
    sentiments) %>% 
    write_csv(metrics_csv)
}
metrics <- read_csv(metrics_csv)

g <- metrics %>%
  ggplot(
    aes(x = date_ym, y = value, color = metric)) +
  geom_line()
ggplotly(g)
```

Interesting! The air quality sentiment (red line) does seem to be inversely correlated with nitrogen dioxide (NO~2~; blue line). The sentiment for clean energy (green line) seems to correlate with satellite NO2 until 2019-03 then remains relatively steady.

## Box plot of satellite & sentiment metrics before/after lockdown

Let's compare averages before and after lockdown March 24, 2020.

```{r}
date_lockdown <- as.Date("2020-03-24")

metrics_ba <- bind_rows(
  metrics %>% 
    filter(date_ym <= date_lockdown) %>% 
    mutate(
      period = "before"),
  metrics %>% 
    filter(date_ym > date_lockdown) %>% 
    mutate(
      period = "after")) %>% 
  mutate(
    period = factor(period, c("before", "after"), ordered = T))
  
# View(metrics_ba)
metrics_ba %>% 
  ggplot() +
  geom_boxplot(aes(x = metric, y = value, fill = metric)) +
  facet_wrap(~period)
```

The sentiments do not look significantly different before and after lockdown, especially relative to the dramatically different NO2 (blue).

## References {-}
