
# Integration

## Time series plot of satellite & sentiment metrics

Let's now integrate data from satellite and sentiment into the same monthly time series plot, first rescaling all metrics -1 to 1 for comparison.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

```{r}
library(here)
library(glue)
library(readr)
library(dplyr)
library(lubridate)
library(scales)
library(ggplot2)
library(plotly)

# paths
loc_name       <- "delhi"
satellite_csv  <- here(glue("data/no2_{loc_name}_ee-chart.csv"))
sentiments_csv <- here(glue("data/sentiments_{loc_name}.csv"))
metrics_csv    <- here(glue("data/metrics_{loc_name}.csv"))

lockdown <- list(
  date = as.Date("2020-03-24"),
  text = "national lockdown") 

if (!file.exists(metrics_csv)){
  
  sentiments <- read_csv(sentiments_csv) %>% 
    rename(date = ym_date) %>% 
    mutate(
      metric    = glue("{query_name}"),
      sentiment = tf_score) %>% 
    select(metric, date, sentiment)
  
  satellite  <- read_csv(satellite_csv) %>% 
    transmute(
      date = mdy(`system:time_start`) %>% 
        floor_date(date, unit = "month"),
      no2  = NO2_column_number_density) %>% 
    group_by(date) %>% 
    summarize(
      no2 = mean(no2, na.rm=T)) %>% 
    mutate(
      metric    = "NO2",
      sentiment = rescale(no2, to = range(sentiments$sentiment))) %>% 
    select(metric, date, no2, sentiment)
  
  bind_rows(
    satellite,
    sentiments) %>% 
    write_csv(metrics_csv)
}
metrics <- read_csv(metrics_csv)

no2 <- metrics %>% 
  filter(metric == "NO2") %>% 
  pull(no2)

g <- metrics %>%
  ggplot(
    aes(x = date, y = sentiment, color = metric)) +
  geom_line() +
  scale_y_continuous(
    sec = sec_axis(
      ~rescale(., range(no2)),
      name = bquote(NO[2]~"(µmol/"~m^2~")"))) +
  theme(legend.position="top")

g_lockdown <- g +
  geom_vline(xintercept = lockdown$date) +
  annotate(
    "text", label = lockdown$text,
    x = lockdown$date, y = 0, angle = 90, vjust = -1)
g_lockdown

ggplotly(g)
```

Interesting! The air quality sentiment (red line) does seem to be inversely correlated with nitrogen dioxide (NO~2~; blue line). The sentiment for clean energy (green line) seems to correlate with satellite NO2 until 2019-03 then remains relatively steady.

## Box plot of satellite & sentiment metrics before/after lockdown

Let's compare averages before and after lockdown March 24, 2020.

```{r}
date_lockdown <- as.Date("2020-03-24")

metrics_ba <- bind_rows(
  metrics %>% 
    filter(date <= date_lockdown) %>% 
    mutate(
      period = "before"),
  metrics %>% 
    filter(date > date_lockdown) %>% 
    mutate(
      period = "after")) %>% 
  mutate(
    period = factor(period, c("before", "after"), ordered = T))
  
g <- metrics_ba %>% 
  ggplot() +
  geom_boxplot(aes(x = metric, y = sentiment, fill = metric)) +
  facet_wrap(~period)
ggplotly(g)
```

The sentiments do not look significantly different before and after lockdown, especially relative to the dramatically different NO2 (blue).

## References {-}
